Effects of gluon kinematics and the Sudakov form factor\\ on the dipole amplitude

### 2023-05-08 Paper/draft.tex
documet update

\usepackage[english]{babel}
\usepackage[a4paper]{geometry}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{breqn}
\usepackage[]{fancyhdr}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{ulem}
%\usepackage{times}
%\usepackage{merriweather}
%\usepackage{draftwatermark}
%\usepackage[backend=biber, style=ieee]{biblatex}

%\addbibresource{draft.bib}

@@

%\newcommand{\comment}[1]{{\color{red}[\textbf{comment:} #1 ]}}

\pagenumbering{arabic}
\numberwithin{equation}{section}
\numberwithin{table}{section}
\numberwithin{figure}{section}

\newcommand{\fww}[0]{\mathcal{F}^{\mathrm{WW}}}
\newcommand{\fdp}[0]{\mathcal{F}^{\mathrm{dipole}}}
\newcommand{\sdp}[0]{\sigma_{\mathrm{dipole}}}
\newcommand{\sdpa}[0]{\sigma_{\mathrm{dipole\,Adj.}}}
\newcommand{\GeV}[0]{\mathrm{GeV}}
\newcommand{\comment}[1]{\texttt{\color{red}#1}}
\newcommand{\pairdot}[2]{ \mathbf{#1}\cdot\mathbf{#2}  }
\author{
T.~Goda,$^1$, K.~Kutak$^{1}$ and S.~Sapeta$^1$\\\,\\
$^1$ 
{\small\it The H.\ Niewodnicza\'nski Institute of Nuclear Physics PAN,}\\ 
{\small\it Radzikowskiego 152, 31-342 Krak\'ow, Poland}\\%\,\\
}	 
%\title{ New parameters for the Golec-Beiernat--W\"usthoff and Bartels--Golec-Biernat--Kowlski saturation models in the $k_T$-facorization formula of $F_2$}
\title{Effects of gluon kinematics and the Sudakov form factor on the GBW and BGK saturation models}
\date{\today}

\begin{document}

@@

%--------------- preprint numbers ---------------------
\vspace{-25em}
\begin{flushright}
  IFJPAN-IV-2023-XX\\
\end{flushright}
\vspace{20em}
%------------------------------------------------------

\begin{abstract}
We investigate effects of exact gluon kinematics on the parameters of the
Golec-Biernat--W\"usthoff, and Bartels--Golec-Biernat--Kowalski models. The
resulting parameters show sizable difference, particularly in the normalization
$\sigma_0$. The resulting models are applied for the dijet process in DIS to
investigate effects of the Sudakov form factor.  The result qualitatively agrees
with that of Ref.~\cite{vanHameren:2021sqc}, except for the difference stemming
from the large-$k_T$ behaviour of the models.
\end{abstract}

@@

@@

  \draw[ snake=coil,segment aspect=1.5,segment length=0.1in](0.5in , -1in)--(0.5in,-0.5in);
  \draw[snake=coil,segment aspect=1.5,segment length=0.1in](-0.5in , -1in)--(-0.5in,-0.5in);
\end{tikzpicture}
}
\caption{Kinematic variables of the gluon contribution to $F_2$. The momentum fraction of gluon is $x/z$ as it has to contribute via the quark box.
}
\label{fig:gluon-kinem}
\end{figure}
Factorization has played central role in QCD, where long-distance effects can be isolated into an object called parton distribution functions (PDFs)~\cite{roberts_1990,collins_2011}. 
The PDFs are not perturbatively calculable but they are process independent~\cite{collins_2011}. That is to say, that one needs to measure them in experiments, but the PDFs obtained one experiment can be used as input in another. A type of experiments called deep inelastic scattering~(DIS), has been one of experiments which largely contributed to the study of the structure of the proton~\cite{roberts_1990,collins_2011}. 
The election-ion collider (EIC)~\cite{NAP25171}, which is currently under construction, will be the next generation of the DIS experiments and expected to uncover more details about the hadron structure.  
As one increases the center of mass energy at a fixed value of the photon virtuality, $Q$, one is probing the region of the small Bjorken scaling variable, $x$.
In the small-$x$ region of DIS, % process   
interactions are dominated by gluons\cite{roberts_1990,collins_2011, Kovchegov:2012mbw}. As a photon does not interact directly with a gluon, it interact through a box of quarks as depicted in Fig.~\ref{fig:gluon-kinem}. 
An object, which is central to the description of this process is called dipole gluon density, $\mathcal{F}(x,k_T^2)$, which is a type of transverse momentum dependent PDFs (TMD PDF)~\cite{Bomhof:2006dp,Dominguez:2010xd,Dominguez:2011wm,Xiao:2017yya}. This appears in the $k_T$-factorization formula~\cite{Catani:1990eg,Dominguez:2010xd,Dominguez:2011wm,Xiao:2017yya}, which we will discuss later, and plays role similar to that of collinear parton distribution function in the collinear factorization.
Much effort has been made to study this object, notably the work of Balistky, Fadin, Kuraev and Lipatov (BFKL)~\cite{Balitsky:1978ic, Kuraev:1977fs} has predicted sharp rise of cross sections with $1/x$.  However, such rise violates unitarity and the Froissart bound~\cite{roberts_1990,collins_2011,Kovchegov:2012mbw,Barone:1993sy}. In \cite{Gribov:1983ivg} it has been recognized that gluon recombination can tame the growth of gluons \cite{Gribov:1983ivg}. The interplay of linear and nonlinear term leads to an phenomenon called gluon saturation. \comment{a line mentioning BK is removed. Is there a particular reason not to mention BK or simply GLR replaced them? }% 
{\color{gray}Similarly, the work of Balitsky and Kovchegov (BK)~\cite{Balitsky:1995ub,Kovchegov:1999yj} incorporated the gluon recombination in the evolution in the dipole formalism, thus successfully incorporating the saturation.}
 The meaning of the Froissart bound is that, asymptotically, the cross section can not grow faster than the size of black-disk (proton with high gluon density) which expands like $\ln(s)$~\cite{Badelek:1992gs}.
Despite the success of such evolution equations, there are several popular phenomenological models owing to their simplicity.

\sout{\comment{Some more general talk like in the previous paper}\\ }
\sout{\comment{EIC} }

In the next section, differences of the $k_T$-factorization and the dipole-factorization are outlined.  In Sec.~\ref{Results}, results of the new fit to HERA data in the $k_T$-factorization formula are presented and discussed. In Sec.~\ref{Dijet process at EIC}, we apply the results to compute cross section for the dijet process in DIS at the Electron-Ion Collider (EIC)~\cite{NAP25171} and make comparisons with Ref.~\cite{vanHameren:2021sqc}.\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The framework}

A notable example is the Golec-Biernat--W\"sthoff saturation model~\cite{Golec-Biernat:1998zce}~(also \cite{Kowalski:2003hm,McLerran:1993ni,Forshaw:1999uf,Iancu:2003ge}). The model was originally proposed in the position space version of the $k_T$-factorization formula, and the object which corresponds to the gluon density of the $k_T$-factorization is called dipole cross section.
In the GBW model, the dipole cross section has the form~\cite{Golec-Biernat:1998zce}
\begin{align}
\sigma_{\mathrm{GBW}}(x,r)&=\sigma_0\left(1-e^{-r^2/R^2_0}\right)&\mathrm{where}&
&R^{-2}_0&=\frac{Q_0^{2}}{4}\left(\frac{x_0}{x}\right)^\lambda.
\end{align}
%\comment{here I use $R_0$ instead of $Q_s$ since for BGK below, $Q_s\neq 1/R_0$. }
The dipole cross section is related to the gluon density by the Fourier transform: 
\begin{equation}
\alpha_s\mathcal{F}(x,k_T^2)=\frac{N_c}{4\pi}\int\frac{d^2\mathbf{r}}{(2\pi)^2}e^{i\mathbf{k}_t\cdot \mathbf{r} }\nabla_{\mathbf{r}}^2\sigma_{\mathrm{dipole}}(x,r).
\label{eq:dipole-gluon}
\end{equation}
%An important feature of the GBW model is that it describes the phenomenon of saturation in a simple form. 
The essence of GBW model is encoded in the $x$-dependent saturation scale $Q_s^2(x)=Q^2_0(x_0/x)^{-\lambda}$, which separates the saturation region and the scaling region.
A further development was made by Bartels, Golec-Biernat and Kowalski~(BGK)~\cite{Bartels:2002cj} which incorporates the DGLAP evolution by modifying the exponent to 
\begin{equation}
R_0^{-2}=\frac{\pi^2\alpha(\mu^2)xg(x,\mu^2)}{3\sigma_0},
\end{equation}
where
\begin{align}
\mu^2&=\frac{C}{r^2}+\mu_0^2 & &\mathrm{and} & xg(x,Q^2_0)&=A_g x^{-\lambda_g}(1-x)^{5.6},
\end{align}
thus improving the description of data at higher $Q^2$.\\
While these models enjoyed much success in phenomenology of DIS, including the diffractive and photo-production processes~\cite{Golec-Biernat:1998zce,Golec-Biernat:1999qor}, as later discussed in detail, derivation of the dipole-factorization formula from the $k_T$-factorization involves with certain simplification and introduction of massive light quarks.

Now, let us introduce $q'\equiv q+x p$. 
Then, one  decomposes $k$ and $\kappa$ (Fig.~\ref{fig:gluon-kinem}) as
\begin{align}
    \kappa&=\alpha p-\beta q'+\kappa_t&&\mathrm{and}& k&=a p- bq'+k_T.
\end{align}
For the contribution depicted in Fig.~\ref{fig:gluon-kinem}, the structure function, $F_2$ for transversely and longitudinally polarized photons, factorize, with a change of variable  ${\boldsymbol{\kappa}'}_t\equiv{\boldsymbol{\kappa}_t}-(1-\beta)\mathbf{k}_t$, in the form~\cite{ Kimber:2001uaa,Kwiecinski:1997ee}
\begin{multline}
	F_2(x,Q^2)=\sum_f e_f^2 \frac{Q^2}{2\pi}\int\frac{dk^2_t}{k_T^2}\int^1_0d\beta\int d{\kappa'}_t\alpha_s(\mu^2) \mathcal{F}(x/z,k_T^2)\Theta(1-x/z)\\
	\left[\left(\beta^2+(1-\beta)^2\right)\left(\frac{I_1}{2\pi}-\frac{I_2}{2\pi}\right)
	+\left(m_f^2+4Q^2\beta^2(1-\beta)^2\right)\left(\frac{I_3}{2\pi}-\frac{I_4}{2\pi}\right)\right],
	\label{eq:angle-integrated}
\end{multline}
where
\begin{equation}
	\frac{1}{z}=1+\frac{{\kappa'}^2_t+m_f^2}{\beta(1-\beta)Q^2}+\frac{k^2_t}{Q^2}.
	\label{eq:z}
\end{equation}
(see Ref.~\cite{ Kimber:2001uaa,Kwiecinski:1997ee} for detail.)
\sout{\color{red} We need an appendix with expressions for $I_i$}
%and

%Hereafter, we will call Eq.~(\ref{eq:angle-integrated}) $k_T$-factorizaton formula.
One should note, that gluons are not probed directly, thus the argument of the gluon density is $x/z$ rather than $x$.
If one, instead, uses $ 1/z=\left(1+4 m_f^2/Q^2\right)$ and assumes that $\mu$ is independent of $\kappa'$ and $k_T$, 
the above formula can be written in the impact parameter space~\cite{Golec-Biernat:1998zce,Nikolaev:1990ja} (i.e. dipole-factorizaton formula)
\begin{equation}
F_2\left(x,Q^2\right)\sim\int^1_0d\beta \int d^2\mathbf{r}\left|\Psi\left(\tilde{x},\beta,Q^2\right)\right|^2\sigma_{\mathrm{dipole}}\left(\tilde{x},r\right),
\label{eq:dipole-factorization}
\end{equation}
where the photon wave function, $\left|\Psi\left(x,\beta,Q^2\right)\right|^2$, describes splitting of the incoming photon into a $q\overline{q}$ pair with light-cone momenta fractions $\beta$ and $1-\beta$ respectively, and the dipole cross section, $\sigma_{\mathrm{dipole}}\left(x,r\right)$, describes the interaction of the colour dipole of the size $r$ with the proton. In Ref.~\cite{Golec-Biernat:1998zce}, $\tilde{x}=x(1+4m_f/Q^2)$ with nonzero vale of $m_f$ was used even for light flavours, which is necessary to partially substitute $k_T$ and ${\kappa'}_T$ in Eq~(\ref{eq:z}). \\

The original GBW and BGK models were fitted in the dipole factorization, Eq.~(\ref{eq:dipole-factorization}). In this study, we use Eq.~(\ref{eq:angle-integrated}), where the gluon density is obtained by evaluating Eq.~(\ref{eq:dipole-gluon}).
Comparing, the GBW and BGK models, one identifies that, the GBW model is effectively the BGK model with $\alpha_s(\mu^2)xg(x,\mu^2)\rightarrow \alpha_s(Q_0^2)xg(x,Q_0^2)$.
For this reason, we assume that $\alpha_s$ in Eq.~(\ref{eq:dipole-gluon}) is constant for the GBW model, and thus explicitly multiply by running coupling;
%
\begin{equation}
\alpha_s(\mu^2)\mathcal{F}(x,k_T^2)\rightarrow \alpha_s(\mu)\frac{\alpha_s\mathcal{F}(x,k_T^2)}{0.2},
\end{equation}
where %0.2 is an arbitrary normalization, and 
%
\comment{Question: alphas=0.2}

\begin{equation}
\alpha_s(\mu^2)=\frac{1}{\frac{11 C_A-2n_f}{12\pi}\log\left(\frac{\mu^2}{\Lambda_{\mathrm{QCD}}^2 }\right)},
\end{equation}
and  $\Lambda_{\mathrm{QCD}}^2=0.09\;\GeV^2$.
The factor 0.2 is an arbitrary normalization and the effect is absorbed by $\sigma_0$ and hence, bears no importance.
(For a more detailed analysis of dipole gluon density from the BGK model see. Ref.~\cite{Luszczak:2022fkf}\footnote{In Ref.~\cite{Luszczak:2022fkf}, the coupling constant was treated differently, and the large $k_T$-region was matched to the derivative of $xg(x)$, while we have used series transformation to accelerate the high-$k_T$ integration. Their treatment is also different from the original BGK paper~\cite{Bartels:2002cj}. As a consequence their gluon density is positive in the large $k_T$-region and the overall large-$k_T$ behaviour is significantly different.}
)
While there is some ambiguity on what the argument of $\alpha_s(\mu^2)$ should be, we follow Ref.~\cite{Kwiecinski:1997ee} and use
\begin{equation}
	\mu^2=k_T^2+{\kappa'}_t^2+m_f^2+\mu^2_0,
\end{equation}
where we add $\mu^2_0=4\;\GeV^2$ in order to freeze the coupling at low scales.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fits to $F_2$ data}
\label{Results}

The GBW and BGK models were fitted to the $F_2$ data from HERA~\cite{Abt:2017nkc}. A numerical program was written with help of CERNLIB~\cite{Kolbig:1972zz}, GSL~\cite{GSL}, CUBA~\cite{Hahn:2004fe} and ROOT~\cite{Brun:1997pa} libraries to evaluate $F_2$. The data were selected to be in the range $0.045\leq Q^2\leq 650\;\mathrm{GeV^2}$, $x<0.01$.
As with the previous fit~\cite{Goda:2022wsc}, the $c$ and $b$ flavours were taken into account with the mass $1.3$ and $4.6\;\mathrm{GeV}$, respectively. As discussed earlier, we take the light quarks to be massless.
The fitting was performed using MnMigrad and MnSimplex of ROOT::Minuit2~\cite{James:2004xla}.

The fits were performed for the following cases
\begin{itemize}
\item GBW model with the fixed coupling in $k_T$-factorization  ($k_T$-GBW),
\item GBW model with the running coupling in $k_T$-factorization (rc-$k_T$-GBW),
\item BGK model in $k_T$-factorization ($k_T$-BGK),
\end{itemize} 
and, as a reference, we provide the following results from Ref.~\cite{Goda:2022wsc};
\begin{itemize}
\item GBW model with massless light quarks in the dipole-factorization ($r$-GBW),
\item GBW model with massive light quarks in the dipole-factorizations ($r$-GBW-massive),
\item  BGK model in  the dipole-factorization ($r$-BGK).
\end{itemize}

>>>>

\usepackage[english]{babel}
\usepackage[a4paper]{geometry}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{breqn}
%\usepackage[]{fancyhdr}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{cite}
%\usepackage{ulem}

@@

%-----------------------------------------------------------------------------
%\pagenumbering{arabic}
%\numberwithin{equation}{section}

%-----------------------------------------------------------------------------
\newcommand{\fww}[0]{\mathcal{F}^{\mathrm{WW}}}
\newcommand{\fdp}[0]{\mathcal{F}^{\mathrm{dipole}}}
\newcommand{\sdp}[0]{\sigma_{\mathrm{dipole}}}
\newcommand{\sdpa}[0]{\sigma_{\mathrm{dipole\,Adj.}}}
\newcommand{\GeV}[0]{\mathrm{GeV}}
\newcommand{\comment}[1]{\texttt{\color{red}#1}}
\newcommand{\commentPending}[1]{\texttt{\color{red!25}#1}}
\newcommand{\pairdot}[2]{ \mathbf{#1}\cdot\mathbf{#2}  }

%-----------------------------------------------------------------------------
\author{
T.~Goda,$^1$, K.~Kutak$^{1}$ and S.~Sapeta$^1$\\\,\\
$^1$ 
{\small\it The H.\ Niewodnicza\'nski Institute of Nuclear Physics PAN,}\\ 
{\small\it Radzikowskiego 152, 31-342 Krak\'ow, Poland}\\
}

\title{Effects of gluon kinematics and the Sudakov form factor\\ on the dipole amplitude}

\date{}

%-----------------------------------------------------------------------------
\begin{document}

@@

%--------------- preprint numbers ---------------------
\vspace{-22em}
\begin{flushright}
  IFJPAN-IV-2023-XX
\end{flushright}
\vspace{17.5em}
%------------------------------------------------------

\begin{abstract}
We investigate effects of exact gluon kinematics on the parameters of the
Golec-Biernat--W\"usthoff, and Bartels--Golec-Biernat--Kowalski saturation
models. The resulting parameters show some differences, particularly, the
normalization of the dipole cross section $\sigma_0$. The refitted models are
used for the dijet production process in DIS to investigate effects of the
Sudakov form factor at Electron Ion Collider energies.  
\end{abstract}

@@

\comment{preprint number (SS)}\\

@@

  \draw[ snake=coil,segment aspect=1.5,segment length=0.1in](0.5in , -1in)--(0.5in,-0.5in);
  \draw[snake=coil,segment aspect=1.5,segment length=0.1in](-0.5in , -1in)--(-0.5in,-0.5in);
\end{tikzpicture}
}
\caption{Kinematic variables of the gluon contribution to $F_2$. The momentum fraction of gluon is $x/z$ as it has to contribute via the quark box.
}
\label{fig:gluon-kinem}
\end{figure}

Factorization of scales plays central role in {\it Quantum Chromodynamics}
(QCD). In particular, within collinear factorization approach, long-distance
effects can be isolated into objects called {\it collinear parton distribution
functions}~(PDFs)~\cite{Roberts:1990ww,Collins:2011zzd}.  The collinear PDFs are not
fully perturbatively calculable but they obey perturbative evolution equations
and are process independent~\cite{Collins:2011zzd}. That is to say that one needs
to measure them in experiments, but the PDFs obtained in one experiment can be
used as input in another.

One of the main type of processes that is suited particularly well to the studies
of proton structure is the {\it deep inelastic
scattering}~(DIS)~\cite{Roberts:1990ww,Collins:2011zzd}. The data from HERA has
enabled us to answer many questions in that domain and largely improved our
picture of interior of a proton. 
It is therefore very exiting that the next generation of a DIS machines,
the election-ion collider~(EIC)~\cite{NAP25171} is making its way and will soon
allow us to uncover more details about the hadron structure.

With an increase of the of the center of mass energy at fixed value of the
photon virtuality, $Q$, one shall probe  the region of the small Bjorken scaling
variable, $x$.  
In that region, dominated by gluons\cite{Roberts:1990ww,Collins:2011zzd,
Kovchegov:2012mbw}, the appropriate theoretical formalism to use is the Color
Glass Condensate~\cite{McLerran:1993ni, Ayala:1995kg, Kovchegov:1996ty,Jalilian-Marian:1996mkd,Iancu:2000hn,Iancu:2001ad,Ferreiro:2001qy,Iancu:2002xk} or, if the density of gluons is still moderate, the
High Energy Factorization also called $k_T$-factorization~\cite{Catani:1990xk,Catani:1990eg,Collins:1991ty,Catani:1993ww,Catani:1993rn,Catani:1994sq}.

In the above frameworks, the interaction of the photon with gluons happens
through a box of quarks, as depicted in Fig.~\ref{fig:gluon-kinem}. An object,
which is central to the description of this process is called the dipole gluon
density, $\fdp(x,k_T^2)$, which is a type of {\it transverse momentum
dependent} PDFs (TMD
PDFs)~\cite{Bomhof:2006dp,Dominguez:2010xd,Dominguez:2011wm,Xiao:2017yya}. 
%
It appears in the $k_T$-factorization
formula~\cite{Catani:1990eg,Dominguez:2010xd,Dominguez:2011wm,Xiao:2017yya},
which we shall discuss in the next section, and plays a role similar to that of
the collinear parton distribution function in the collinear factorization.

Much effort has been made
to study this object, notably the work of Balitsky, Fadin, Kuraev and Lipatov
(BFKL)~\cite{Balitsky:1978ic, Kuraev:1977fs} has predicted sharp rise of cross
sections with $1/x$.  However, such rise violates unitarity and the Froissart
bound~\cite{Roberts:1990ww,Collins:2011zzd,Kovchegov:2012mbw,Barone:1993sy}. In
\cite{Gribov:1983ivg} it has been recognized that gluon recombination can tame
the growth of gluons \cite{Gribov:1983ivg}. The interplay of linear and
nonlinear terms leads to a phenomenon called {\it gluon saturation} and
corresponding evolution equations are known as
Balitsky-Kovchegov~(BK)~\cite{Balitsky:1995ub,Kovchegov:1999yj} equation or
JIMWLK\cite{Kovner:1999bj,Kovner:2000pt,Iancu:2000hn,JalilianMarian:1997dw,JalilianMarian:1997gr,JalilianMarian:1997jx} equation.

Due to its clear picture, one often uses a position space counter part of the gluon density, called {\it dipole cross section}. Despite the success of the aforementioned evolution equations, there exists a number of phenomenological models of the dipole cross section, which are popular due to to their simplicity. 
\commentPending{More about saturation models}
A notable example of a dipole saturation model is the one proposed by
Golec-Biernat and Wu\"sthoff (GBW)~\cite{Golec-Biernat:1998zce} and the one proposed by
Bartels, Golec-Biernat and Kowalski (BGK)~\cite{Bartels:2002cj}. We will use these two models in this study. There are other models, with different considerations, such as impact parameter dependence~({\it cf.}
\cite{Kowalski:2003hm,McLerran:1993ni,Forshaw:1999uf,Iancu:2003ge}).

\commentPending{Section about what we did: fits of saturation models with exact
kinematics, use of those results + Sudakov for EIC predictions.}

In the present study, we fit the GBW~\cite{Golec-Biernat:1998zce} and BGK~\cite{Bartels:2002cj} models in the $k_T$ factorization formula of $F_2$, instead of the original dipole factorization formula. As discussed below, this allows us to fit the model with exact gluon kinematics. 
Since inclusive observables do not reveal the $k_T$ dependence well, we also consider a less inclusive observable which is more sensitive to the $k_T$ dependence of the gluon distribution.
The resulting models are put to the test by applying them to the predictions of dijet DIS at EIC, including effects of the Sudakov form factor, which resums logarithms of the small transverse momentum.

The paper is organized as follows.  In the next section, we present theoretical
framework which we used and, in particular, outline differences of the
$k_T$-factorization and the dipole-factorization.  In
Sec.~\ref{sec:F2fits}, results of the new fits to HERA data in the
$k_T$-factorization formula are presented. 
In Sec.~\ref{sec:dijetsEIC}, we apply the results from the previous section to
compute distributions for dijet production in DIS at the Electron-Ion Collider
(EIC)~\cite{NAP25171} and make comparisons to our earlier study of Ref.~\cite{vanHameren:2021sqc}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The framework}
\commentPending{Missing a general factorization formula}
In general, a factorization formula has the form
\begin{equation}
    \sigma=H \otimes \phi,
\end{equation}
where $\otimes$ denotes convolution. In this form one achieves two things. Firstly, one can confine long distance effects in one of the factors (let $\phi$ be such an object), and leaves $H$ as an perturbatively calculable object. Secondly, one may factorize out a (to some extent) process independent object.  
In the present study, the dipole cross section and the gluon densities both  contain nonperturbative part and are, to an extent, process independent~\footnote{They are not universal, however they may appear in different processes. (see for example Refs.~\cite{Dominguez:2011wm,Kotko:2015ura})}.
In this section, we will outline two factorizations; the $k_T$ factorization and the dipole factorization. \\
As it was mentioned earlier, we use the GBW and BGK models.  
The models were originally formulated in the position-space version of the
$k_T$-factorization formula. %, and the object which corresponds to the gluon
%density of the $k_T$-factorization is called the dipole cross section. 
In the GBW model, the dipole cross section has the form~\cite{Golec-Biernat:1998zce}
\begin{align}
\label{eq:gbw-dipole}
\sigma_{\mathrm{GBW}}(x,r)&=\sigma_0\left(1-e^{-r^2/R^2_0}\right)&\mathrm{where}&
&R^{-2}_0&=\frac{Q_0^{2}}{4}\left(\frac{x_0}{x}\right)^\lambda.
\end{align}
%
The dipole cross section is related to the gluon density by the Fourier
transform
%
\begin{equation}
\alpha_s\fdp(x,k_T^2)=\frac{N_c}{4\pi}\int\frac{d^2\mathbf{r}}{(2\pi)^2}e^{i\mathbf{k}_t\cdot \mathbf{r} }\nabla_{\mathbf{r}}^2\sigma_{\mathrm{dipole}}(x,r).
\label{eq:dipole-gluon}
\end{equation}
%
The essence of the GBW model is encoded in the $x$-dependent saturation scale
$Q_s^2(x)=Q^2_0(x_0/x)^{-\lambda}$, which separates the saturation region and
the scaling region.  A extension of the above model was proposed by Bartels, Golec-Biernat
and Kowalski~(BGK)~\cite{Bartels:2002cj} who incorporated the DGLAP evolution
in the GBW cross section~(\ref{eq:gbw-dipole}) by modifying the exponent to 
%
\begin{equation}
R_0^{-2}=\frac{\pi^2\alpha(\mu^2)xg(x,\mu^2)}{3\sigma_0},
\end{equation}
where
\begin{align}
\mu^2&=\frac{C}{r^2}+\mu_0^2 & &\mathrm{and} & xg(x,Q^2_0)&=A_g x^{-\lambda_g}(1-x)^{5.6},
\end{align}
%
thus improving the description of data at higher $Q^2$.\\

While these models enjoyed much success in phenomenology of DIS, including the
diffractive and photo-production
processes~\cite{Golec-Biernat:1998zce,Golec-Biernat:1999qor}, it is worth
mentioning that they both use certain kinematic approximation which follows from
subtle differences in the dipole factorization and the $k_T$-factorization.  In
the following, we will investigate effects of these approximations.

Firstly, let us outline the factorization formulas.
With $q'\equiv q+x p$, one decomposes $k$ and $\kappa$, defined in
Fig.~\ref{fig:gluon-kinem}, as
%
\begin{align}
    \kappa&=\alpha p-\beta q'+\kappa_t&&\mathrm{and}& k&=a p- bq'+k_T.
\end{align}
For the contribution depicted in Fig.~\ref{fig:gluon-kinem}, the structure function, $F_2$ for transversely and longitudinally polarized photons, factorize, with a change of variable  ${\boldsymbol{\kappa}'}_t\equiv{\boldsymbol{\kappa}_t}-(1-\beta)\mathbf{k}_t$, in the form~\cite{ Kimber:2001uaa,Kwiecinski:1997ee}
\begin{multline}
	F_2(x,Q^2)=\sum_f e_f^2 \frac{Q^2}{2\pi}\int\frac{dk^2_t}{k_T^2}\int^1_0d\beta\int d{\kappa'}_t\alpha_s(\mu^2) \fdp(x/z,k_T^2)\Theta(1-x/z)\\
	\left[\left(\beta^2+(1-\beta)^2\right)\left(\frac{I_1}{2\pi}-\frac{I_2}{2\pi}\right)
	+\left(m_f^2+4Q^2\beta^2(1-\beta)^2\right)\left(\frac{I_3}{2\pi}-\frac{I_4}{2\pi}\right)\right],
	\label{eq:angle-integrated}
\end{multline}
where
\begin{equation}
	\frac{1}{z}=1+\frac{{\kappa'}^2_t+m_f^2}{\beta(1-\beta)Q^2}+\frac{k^2_t}{Q^2}\,.
	\label{eq:z}
\end{equation}

One should note, that gluons are not probed directly, thus the argument of the
gluon density is $x/z$ rather than $x$.  If one, instead, uses $ 1/z=1+4
m_f^2/Q^2$ and assumes that $\mu$ is independent of $\kappa'$ and $k_T$, the
above formula can be written in the impact parameter
space~\cite{Golec-Biernat:1998zce,Nikolaev:1990ja} ({\it i.e.} as a
dipole-factorization formula)
%
\begin{equation}
  F_2\left(x,Q^2\right)=\sum_f e_f^2 \frac{Q^2}{4\pi^2} \int^1_0d\beta \int \frac{d^2\mathbf{r}}{(2\pi)^2} \left|\Psi\left(\tilde{x},\beta,Q^2\right)\right|^2\sigma_{\mathrm{dipole}}\left(\tilde{x},r\right),
  \label{eq:dipole-factorization}
\end{equation}

where the photon wave function, $\left|\Psi\left(x,\beta,Q^2\right)\right|^2$,
describes splitting of the incoming photon into a $q\overline{q}$ pair with
light-cone momenta fractions $\beta$ and $1-\beta$ respectively, and the dipole
cross section, $\sigma_{\mathrm{dipole}}\left(\tilde x,r\right)$, describes the
interaction of the colour dipole of the size $r$ with the proton. In
Ref.~\cite{Golec-Biernat:1998zce}, $\tilde{x}=x(1+4m_f/Q^2)$ with nonzero value
of $m_f$ was used even for light flavours, which is necessary to partially
substitute $k_T$ and ${\kappa'}_T$ in Eq.~(\ref{eq:z}).
%The original GBW and BGK models were fitted in the dipole factorization, Eq.~(\ref{eq:dipole-factorization}), while 
In the present study, we use Eq.~(\ref{eq:angle-integrated}) to fit the models, where the gluon density is obtained by evaluating Eq.~(\ref{eq:dipole-gluon}).
%Comparing, the GBW and BGK models, one identifies that, the GBW model is effectively the BGK model with 
%\begin{equation}
%\begin{split}
%    \alpha_s(\mu^2)xg(x,\mu^2)\rightarrow &\alpha_s(Q_0^2)xg(x,Q_0^2)\\
%    &=\alpha_s x^{-\lambda}(1-x)^{5.6}.
%\end{split}
%\end{equation}
\commentPending{??}{\color{blue}response: the argument on why we consider $\alpha$ as a constant is removed, as it is not a solid argument.}
%For this reason,
In order to account for the running coupling in Eq.~(\ref{eq:angle-integrated}), 
we assume that $\alpha_s$ in Eq.~(\ref{eq:dipole-gluon}) is
constant for the GBW model, and thus explicitly multiply by the running coupling
%
\begin{equation}
\alpha_s(\mu^2)\fdp(x,k_T^2)\rightarrow \alpha_s(\mu)\frac{\alpha_s\fdp(x,k_T^2)}{0.2},
\end{equation}
where
%
\begin{equation}
\alpha_s(\mu^2)=\frac{1}{\frac{11 C_A-2n_f}{12\pi}\log\left(\frac{\mu^2}{\Lambda_{\mathrm{QCD}}^2 }\right)},
\end{equation}
and  $\Lambda_{\mathrm{QCD}}^2=0.09\;\GeV^2$.

The factor 0.2 is an arbitrary normalization whose effect is absorbed by $\sigma_0$ and hence, bears no importance.
(For a more detailed analysis of the dipole gluon density from the BGK model see Ref.~\cite{Luszczak:2022fkf}\footnote{In Ref.~\cite{Luszczak:2022fkf}, the coupling constant was treated differently, and the large $k_T$-region was matched to the derivative of $xg(x)$, while we have used series transformation to accelerate the high-$k_T$ integration. Their treatment is also different from the original BGK paper~\cite{Bartels:2002cj}. As a consequence their gluon density is positive in the large $k_T$-region and the overall large-$k_T$ behaviour is significantly different.}
)
While there is some ambiguity on what the argument of $\alpha_s(\mu^2)$ should be, we follow Ref.~\cite{Kwiecinski:1997ee} and use
\begin{equation}
	\mu^2=k_T^2+{\kappa'}_t^2+m_f^2+\mu^2_0,
\end{equation}
where we add $\mu^2_0=4\;\GeV^2$ in order to freeze the coupling at low scales.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fits to $F_2$ data}
\label{sec:F2fits}

We fitted the GBW and BGK models to the $F_2$ data from
HERA~\cite{Abt:2017nkc}. A numerical program was written with help of
CERNLIB~(DPSIPG)~\cite{Kolbig:1972zz}, GSL~\cite{GSL}, CUBA~\cite{Hahn:2004fe}
and ROOT~\cite{Brun:1997pa} libraries to evaluate $F_2$. The data were selected
to be in the range $0.045\leq Q^2\leq 650\;\mathrm{GeV^2}$, $x<0.01$.  As with
the previous fit~\cite{Goda:2022wsc}, the $c$ and $b$ flavours were taken into
account with the mass $1.3$ and $4.6\;\mathrm{GeV}$, respectively. As discussed
earlier, we take the light quarks to be massless.  The fitting was performed
using MnMigrad and MnSimplex of ROOT::Minuit2~\cite{James:2004xla}.

We studied the following cases
%
\begin{itemize}
\item GBW model with the fixed coupling in $k_T$-factorization  ($k_T$-GBW),
\item GBW model with the running coupling in $k_T$-factorization (rc-$k_T$-GBW),
\item BGK model in $k_T$-factorization ($k_T$-BGK),
\end{itemize} 
%
and, as a reference, we provide the following results from
Ref.~\cite{Goda:2022wsc};
%
\begin{itemize}
\item GBW model with massless light quarks in the dipole-factorization ($r$-GBW),
\item GBW model with massive light quarks in the dipole-factorization ($r$-GBW-massive),
\item  BGK model in  the dipole-factorization ($r$-BGK).
\end{itemize}
